---
title: "Baseball project"
author: "Mari Osmundson"
date: "12/9/2019"
output: word_document
---

Main file: game_log. This file contains data on baseball games played from 1871 to 2016.
  Definitions:
    putouts: the number of times a defensive player got an offensive player out.
    def_pos: Stands for defensive position. Indicates which defense position the player takes.
      Position list follows:
        1. Pitcher
        2. Catcher
        3. First base
        4. Second base
        5. Third base
        6. Shortstop
        7. Left field
        8. Center field
        9. Right field
    The player's number does not necessarily correspond to their defensive position.
Helper files: 
  1) person_codes: contains the IDs, names, and start date for players, coaches, and umpires referenced in the game_log. Primary key is id. Various correlations in the game_log: _umpire_id, _manager_id, _player_id; _pitcher_name, _manager_name, _umpire_name.
  2) park_codes: Contains name, location, league, and usage dates. Primary key is park_id, which corresponds to park_id in the game_log file.
  3) team_codes: Lists the team information, such as league and location. Primary key is team_id, which is contained in game_log/v_name and game_log/h_name. Following is a guide to the league abbreviations.
      
      AA = American Association (1882-1891)
      AL = American League (est. 1901)
      FL = Federal League (1914-1915)
      NL = National League (est. 1876)
      PL = Player's League (1890)
      UA = Union Association (1884)

      
      
I imported game_log and its helper files, then created a database called bb.db using SQLite. bb.db has game_log, person_codes, park_codes, and team_codes as its four tables.

```{r eval=FALSE}
bb.db <- dbConnect(SQLite(), "bb.db")
```

```{r eval=FALSE}
dbWriteTable(bb.db, "person_codes", person_codes, header = TRUE)
dbWriteTable(bb.db, "park_codes", park_codes, header = TRUE)
dbWriteTable(bb.db, "team_codes", team_codes, header = TRUE)
dbWriteTable(bb.db, "game_log", game_log, header = TRUE)
dbWriteTable(bb.db, "appearance_type1", appearance_type1, header = TRUE)
```

I added the game_id column to game_log.
```{r eval=FALSE}
query1 <- "ALTER TABLE game_log
ADD COLUMN game_id TEXT"
dbExecute(bb.db, query1)

dbExecute(bb.db, "UPDATE game_log
SET game_id = h_name || date || number_of_game
WHERE game_id IS NULL")
```

```{r echo=FALSE, eval=FALSE}
query <-"SELECT game_id, h_name, date, number_of_game FROM game_log
LIMIT 5"
result <- dbGetQuery(bb.db, query)
```

After studying the four files, I noticed the following:
game_log: 
  v_league and h_league have no entries and are irrelevant since this data is contained in team_codes.
  The player, coach, umpire, manager names are also contained in the person_codes file and do not need to be in the game_log file. They are attributes of the respective IDs of each person, not of the game they participated in.
  The player positions could be stacked, so that one row consists of a player's defensive and offensive positions.
  
  
I began creating different tables based on the schema I designed in dbdesigner.net, beginning with the person table.
```{r eval=FALSE}
create_table1 <- "CREATE TABLE person (
  id TEXT PRIMARY KEY,
  last_name TEXT,
  first_name TEXT
)"
dbExecute(bb.db, create_table1)

add_columns_table1 <-
"INSERT INTO person
SELECT id, last, first FROM person_codes"
dbExecute(bb.db, add_columns_table1)

query3 <- "SELECT * FROM person
LIMIT 5"
result3 <- dbGetQuery(bb.db, query3)
```

Added league table.
```{r eval=FALSE}
create_league_table <- "CREATE TABLE league (
  league_id TEXT PRIMARY KEY,
  league_name TEXT
)"
dbExecute(bb.db, create_league_table)

add_col_league <- "INSERT INTO league
VALUES ('AA', 'American Association'),
('AL', 'American League'),
('FL', 'Federal League'),
('NL', 'National League'),
('PL', 'Players League'),
('UA', 'Union Association')";
dbExecute(bb.db, add_col_league)

query4 <- "SELECT * FROM league"
result4 <- dbGetQuery(bb.db, query4)
```

Added appearance_type table.
```{r eval=FALSE}
create_app_type_table <- "CREATE TABLE appearance_type (
  appearance_type_id TEXT PRIMARY KEY,
  name TEXT,
  category TEXT
)"
dbExecute(bb.db, create_app_type_table)

add_app_columns <- "INSERT INTO appearance_type
SELECT * FROM appearance_type1";
dbExecute(bb.db, add_app_columns)

query5 <- "SELECT * FROM appearance_type
LIMIT 5"
result5 <- dbGetQuery(bb.db, query5)
```

Added team_info table.
```{r eval=FALSE}
create_team_info <- "CREATE TABLE team_info (
  team_id TEXT PRIMARY KEY,
  league_id TEXT,
  city TEXT,
  nickname TEXT,
  franch_id TEXT,
  FOREIGN KEY (league_id) REFERENCES league(league_id)
)";
dbExecute(bb.db, create_team_info)

add_team_cols <- "INSERT OR IGNORE INTO team_info
SELECT team_id, league, city, nickname, franch_id FROM team_codes";
dbExecute(bb.db, add_team_cols)

query7 <- "SELECT * from team_info
LIMIT 5"
result7 <- dbGetQuery(bb.db, query7)
```

Added game table.
```{r eval=FALSE}
create_game_tbl <- "CREATE TABLE game (
  game_id TEXT PRIMARY KEY,
  date TEXT,
  number_of_game INTEGER,
  day_of_week TEXT,
  park_id TEXT,
  attendance INTEGER,
  length_minutes INTEGER,
  completion TEXT,
  forfeit TEXT,
  protest TEXT,
  additional_info TEXT,
  acquisition_info TEXT,
  FOREIGN KEY (park_id) REFERENCES park_codes(park_id)
)";
dbExecute(bb.db, create_game_tbl)

add_game_cols <- "INSERT INTO game
SELECT 
  game_id, 
  date,
  number_of_game,
  day_of_week,
  park_id,
  attendance,
  length_minutes,
  completion,
  forefeit,
  protest,
  additional_info,
  acquisition_info 
FROM game_log";
dbExecute(bb.db, add_game_cols)

query8 <- "SELECT game_id, date, number_of_game FROM game
LIMIT 5"
result8 <- dbGetQuery(bb.db, query8)
```

Added team_stats table.
```{r eval=FALSE}
create_team_stats <- "CREATE TABLE team_stats (
  team_id TEXT,
  game_id TEXT,
  home BOOLEAN,
  score INTEGER,
  line_score TEXT,
  at_bats INTEGER,
  hits INTEGER,
  doubles INTEGER,
  triples INTEGER,
  homeruns INTEGER,
  rbi INTEGER,
  sacrifice_hits INTEGER,
  sacrifice_flies INTEGER,
  hit_by_pitch INTEGER,
  walks INTEGER,
  intentional_walks INTEGER,
  strikeouts INTEGER,
  stolen_bases INTEGER,
  pitchers_used INTEGER,
  individual_earned_runs INTEGER,
  team_earned_runs INTEGER,
  putouts INTEGER,
  assists INTEGER,
  errors INTEGER,
  double_plays INTEGER,
  triple_plays INTEGER,
  PRIMARY KEY (team_id, game_id),
  FOREIGN KEY (team_id) REFERENCES team(team_id),
  FOREIGN KEY (game_id) REFERENCES game(game_id)
)";
dbExecute(bb.db, create_team_stats)

create_ts_cols <- "INSERT INTO team_stats
SELECT
  h_name,
  game_id,
  1 AS home,
  h_score,
  h_line_score,
  h_at_bats,
  h_hits,
  h_doubles,
  h_triples,
  h_homeruns,
  h_rbi,
  h_sacrifice_hits,
  h_sacrifice_flies,
  h_hit_by_pitch,
  h_walks,
  h_intentional_walks,
  h_strikeouts,
  h_stolen_bases,
  h_pitchers_used,
  h_individual_earned_runs,
  h_team_earned_runs,
  h_putouts,
  h_assists,
  h_errors,
  h_double_plays,
  h_triple_plays
FROM game_log

UNION

SELECT
  v_name,
  game_id,
  0 AS home,
  v_score,
  v_line_score,
  v_at_bats,
  v_hits,
  v_doubles,
  v_triples,
  v_homeruns,
  v_rbi,
  v_sacrifice_hits,
  v_sacrifice_flies,
  v_hit_by_pitch,
  v_walks,
  v_intentional_walks,
  v_strikeouts,
  v_stolen_bases,
  v_pitchers_used,
  v_individual_earned_runs,
  v_team_earned_runs,
  v_putouts,
  v_assists,
  v_errors,
  v_double_plays,
  v_triple_plays
FROM game_log";
dbExecute(bb.db, create_ts_cols)

query9 <- "SELECT * FROM team_stats";
result9 <- dbGetQuery(bb.db, query9)
```

Added person_appearance table.
```{r eval=FALSE}
create_pers_app <- "CREATE TABLE person_appearance (
  appearance_id INTEGER PRIMARY KEY,
  person_id TEXT,
  team_id TEXT,
  game_id TEXT,
  appearance_type_id TEXT
)";
dbExecute(bb.db, create_pers_app)

create_pers_app_t <- "INSERT INTO person_appearance (
    game_id,
    team_id,
    person_id,
    appearance_type_id
)
    SELECT
        game_id,
        NULL,
        lf_umpire_id,
        'ULF'
    FROM game_log
    WHERE lf_umpire_id IS NOT NULL

UNION

    SELECT
        game_id,
        NULL,
        rf_umpire_id,
        'URF'
    FROM game_log
    WHERE rf_umpire_id IS NOT NULL

UNION

    SELECT
        game_id,
        v_name,
        v_manager_id,
        'MM'
    FROM game_log
    WHERE v_manager_id IS NOT NULL

UNION

    SELECT
        game_id,
        h_name,
        h_manager_id,
        'MM'
    FROM game_log
    WHERE h_manager_id IS NOT NULL

UNION

    SELECT
        game_id,
        CASE
            WHEN h_score > v_score THEN h_name
            ELSE v_name
            END,
        winning_pitcher_id,
        'AWP'
    FROM game_log
    WHERE winning_pitcher_id IS NOT NULL

UNION

    SELECT
        game_id,
        CASE
            WHEN h_score < v_score THEN h_name
            ELSE v_name
            END,
        losing_pitcher_id,
        'ALP'
    FROM game_log
    WHERE losing_pitcher_id IS NOT NULL

UNION

    SELECT
        game_id,
        CASE
            WHEN h_score > v_score THEN h_name
            ELSE v_name
            END,
        saving_pitcher_id,
        'ASP'
    FROM game_log
    WHERE saving_pitcher_id IS NOT NULL

UNION

    SELECT
        game_id,
        CASE
            WHEN h_score > v_score THEN h_name
            ELSE v_name
            END,
        winning_rbi_batter_id,
        'AWB'
    FROM game_log
    WHERE winning_rbi_batter_id IS NOT NULL

UNION

    SELECT
        game_id,
        v_name,
        v_starting_pitcher_id,
        'PSP'
    FROM game_log
    WHERE v_starting_pitcher_id IS NOT NULL

UNION

    SELECT
        game_id,
        h_name,
        h_starting_pitcher_id,
        'PSP'
    FROM game_log
    WHERE h_starting_pitcher_id IS NOT NULL";
dbExecute(bb.db, create_pers_app_t)
```

```{r}
dbDisconnect(bb.db)
```





